---
title: "Reppert-IAF-606-Final-Part-3-Random-Forest"
author: "Jeanne Reppert"
date: "4/27/2020"
output: html_document
---
**Set up and read in file from part 1**
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/thepi/OneDrive/Documents/UNCG/IAF 606/Final')

dd <- read.csv(file="cleaned_diabetic_data.csv", header=T, na.strings = "?")
```

**Load libraries for random forest and analysis**
```{r, echo=FALSE}
library(randomForest)
library(caTools)
library(caret)
library(e1071)
```

**drop columns that will not be used for random forest**
```{r}
dd = subset(dd, select = -c(X, ID, Pat_nbr))
```


**reorder columns so that target column is last)
```{r}
dd = dd[c(1,2,3,4,5,6,7,8, 9, 11, 10)]
```


**Create test and train sets and check set dimensions**
```{r}
set.seed(123)
sample = sample.split(dd$Readmitted, SplitRatio = .70)
train = subset(dd, sample == TRUE)
test  = subset(dd, sample == FALSE)
dim(train)
dim(test)
```

## Model 1 ##

**Create first model with all variables included, set cutoff at .10**
```{r}
set.seed(123)
rf1 <- randomForest(Readmitted ~ ., data=train, importance = TRUE, cutoff = c(0.9,0.1))
```

**results of model 1 with training set**
```{r}
rf1
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction <-predict(rf1,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm1 = confusionMatrix(prediction, test$Readmitted, mode = "everything")
cm1
```
**Investigate variable importances**
```{r}
var1 <- varImp(rf1)
var1
```

**Plot variable importances**
```{r}
varImpPlot(rf1)
```

**Calculate Youden's Index**
```{r}
cm1$byClass[1]
cm1$byClass[2]
J1 <- 0.8824237  + 0.1448276   - 1
cat('J1:', J1)
```


## Model 2 ##

**Create second model with all variables except age, set cutoff at .10**
```{r}
set.seed(123)
rf2 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric + 
          Race + HbA1c + Admit + Gender + Discharge, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

**Model 2 training results, error rate increased**
```{r}
rf2
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction2 <-predict(rf2,test[-11])
```

**Create confusion matrix using predictions, accuracy decreased but**
**number of readmissions correctly predicted increased**
```{r}
cm2 = confusionMatrix(prediction2, test$Readmitted, mode = "everything")
cm2
```
**Investigate variable importances**
```{r}
var2 <- varImp(rf2)
var2
```

**Plot variable importances**
```{r}
varImpPlot(rf2)
```

**Calculate Youden's Index**
```{r}
cm2$byClass[1]
cm2$byClass[2]
J2 <- 0.8580922 + 0.1846154  - 1
cat('J2:', J2)
```


## Model 3 ##

**Create model with all variables except age and discharge, set cutoff at .10**
**Removing discharge appears to significantly reduce the models effectiveness**
```{r}
set.seed(123)
rf3 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric +
              Race + HbA1c + Admit + Gender, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

**Model 3 output**
**Accuracy increased but fewer readmitted predicted**
```{r}
rf3
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction3 <-predict(rf3,test[-11])
```

**Create confusion matrix using predictions**
**Accuracy rate increased but correctly predicted readmitted**
**values went down.**
```{r}
cm3 = confusionMatrix(prediction3, test$Readmitted, mode = "everything")
cm3
```
**Investigate variable importances**
```{r}
var3 <- varImp(rf3)
var3
```

**Plot variable importances**
```{r}
varImpPlot(rf3)
```

**Calculate Youden's Index**
**Very low index**
```{r}
cm3$byClass[1]
cm3$byClass[2]
J3 <- 0.9510753 + 0.05676393  - 1
cat('J3:', J3)
```
`


## Model 4 ##

**Create model with all variables except age, and gender (adding discharge back in), set cutoff at .10**
```{r}
set.seed(123)
rf4 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric +
              Race + HbA1c + Admit + Discharge, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

**Model 4 training output**
**Relatively low error rate**
**
```{r}
rf4
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction4 <-predict(rf4,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm4 = confusionMatrix(prediction4, test$Readmitted, mode = "everything")
cm4
```
**Investigate variable importances**
```{r}
var4 <- varImp(rf4)
var4
```

**Plot variable importances**
```{r}
varImpPlot(rf4)
```

**Calculate Youden's Index**
**Almost 0**
```{r}
cm4$byClass[1]
cm4$byClass[2]
J4 <- 0.9530637   + 0.05464191  - 1
cat('J4:', J4)
```



## Model 5 ##

**Create model with all variables except age, gender and admit,**
**set cutoff at .10**
```{r}
set.seed(123)
rf5 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric + Discharge + 
              Race + HbA1c, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

**Model 5 output**
```{r}
rf5
```


**Model 5 output**
```{r}
rf5
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction5 <-predict(rf5,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm5 = confusionMatrix(prediction5, test$Readmitted, mode = "everything")
cm5
```
**Investigate variable importances**
```{r}
var5 <- varImp(rf5)
var5
```

**Plot variable importances**
```{r}
varImpPlot(rf5)
```

**Calculate Youden's Index**
**Too low**
```{r}
cm5$byClass[1]
cm5$byClass[2]
J5 <- 0.9588195   + 0.04244032 - 1
cat('J5:', J5)
```


## Model 6 ##

**Create model with all variables except age, gender, admit,**
**and HbA1c, set cutoff at .10**
```{r}
set.seed(123)
rf6 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric + Discharge + 
              Race, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```
 **Model 6 output**
```{r}
rf6
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction6 <-predict(rf6,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm6 = confusionMatrix(prediction6, test$Readmitted, mode = "everything")
cm6
```
**Investigate variable importances**
```{r}
var6 <- varImp(rf6)
var6
```

**Plot variable importances**
```{r}
varImpPlot(rf6)
```

**Calculate Youden's Index**
**Still too low**
```{r}
cm6$byClass[1]
cm6$byClass[2]
J6 <- 0.9678719 +  0.03766578  - 1
cat('J6:', J6)
```

## Model 7 ##

**Create model with all variables except age, gender,**
**admit, HbA1c, and Race, set cutoff at .10**
```{r}
set.seed(123)
rf7 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric + Discharge,
              data=train, importance = TRUE,
              cutoff = c(0.9,0.1))
```

**Model 7 training output**
```{r}
rf7
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction7 <-predict(rf7,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm7 = confusionMatrix(prediction7, test$Readmitted, mode = "everything")
cm7
```
**Investigate variable importances**
```{r}
var7 <- varImp(rf7)
var7
```

**Plot variable importances**
```{r}
varImpPlot(rf7)
```

**Calculate Youden's Index**
**Still too low**
```{r}
cm7$byClass[1]
cm7$byClass[2]
J7 <- 0.9772906 + 0.02493369  - 1
cat('J7:', J7)
```


## Model 8 ##

**Create model with Time_hosp, Diag and Med_spec, Discharge, set cutoff at .10**
```{r}
set.seed(123)
rf8 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Discharge, 
                    importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

**Model 8 training output**
```{r}
rf8
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction8 <-predict(rf8,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm8 = confusionMatrix(prediction8, test$Readmitted, mode = "everything")
cm8
```
**Investigate variable importances**
```{r}
var8 <- varImp(rf8)
var8
```

**Plot variable importances**
```{r}
varImpPlot(rf8)
```

**Calculate Youden's Index**
```{r}
cm8$byClass[1]
cm8$byClass[2]
J8 <- 0.9955 + 0.003713528   - 1
cat('J8:', J8)
```



## Model 9 ##

**Create model with Time_hosp, Admit, Discharge, HbA1c and Diag, set cutoff at .10**
```{r}
set.seed(123)
rf9 <- randomForest(Readmitted ~ Time_hosp + Diag + Discharge + Admit + HbA1c,
                    importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

```{r}
rf9
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction9 <-predict(rf9,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm9 = confusionMatrix(prediction9, test$Readmitted, mode = "everything")
cm9
```
**Investigate variable importances**
```{r}
var9 <- varImp(rf9)
var9
```

**Plot variable importances**
```{r}
varImpPlot(rf9)
```

**Calculate Youden's Index**
**Still too low**
```{r}
cm9$byClass[1]
cm9$byClass[2]
J9 <- 0.995657 + 0.005835544  - 1
cat('J9:', J9)
```



## Model 10 ##

**Create model with all variables except age_numeric, set cutoff at .10**
```{r}
set.seed(123)
rf10 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age +
              Race + HbA1c + Admit + Gender + Discharge, data=train,
              importance = TRUE, cutoff = c(0.9,0.1))
```

**Model 10 training output**
**This model looks more promising.**
```{r}
rf10
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction10 <-predict(rf10,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm10 = confusionMatrix(prediction10, test$Readmitted, mode = "everything")
cm10
```
**Investigate variable importances**
```{r}
var10 <- varImp(rf10)
var10
```

**Plot variable importances**
```{r}
varImpPlot(rf10)
```

**Calculate Youden's Index**
**Not as good as model with age_numeric (model 1)**
**instead of age categorical.**
```{r}
cm10$byClass[1]
cm10$byClass[2]
J10 <- 0.9135053 + 0.1098143 - 1
cat('J10:', J10)
```

## Model 11 ##

**Create model with Time_hosp, Diag, Med_spec, Age_numeric and HbA1c, set cutoff at .10**
```{r}
set.seed(123)
rf11 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric +
              HbA1c, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

```{r}
rf11
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction11 <-predict(rf11,test)
```

**Create confusion matrix using predictions**
```{r}
cm11 = confusionMatrix(prediction11, test$Readmitted, mode = "everything")
cm11
```
**Investigate variable importances**
```{r}
var11 <- varImp(rf11)
var11
```

**Plot variable importances**
```{r}
varImpPlot(rf11)
```

**Calculate Youden's Index**
**Negative?**
```{r}
cm11$byClass[1]
cm11$byClass[2]
J11 <- 0.9652033 + 0.02970822 - 1
cat('J11:', J11)
```


## Model 12 ##

**Create model with Time_hosp, Diag, Med_spec, Age_numeric, Admit,**
**Discharge and HbA1c, set cutoff at .10**
```{r}
set.seed(123)
rf12 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric +
              Discharge + Admit + HbA1c, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```


```{r}
rf12
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction12 <-predict(rf12,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm12 = confusionMatrix(prediction12, test$Readmitted, mode = "everything")
cm12
```
**Investigate variable importances**
```{r}
var12 <- varImp(rf12)
var12
```

**Plot variable importances**
```{r}
varImpPlot(rf12)
```

**Calculate Youden's Index**
```{r}
cm12$byClass[1]
cm12$byClass[2]
J12 <- 0.9614358 + 0.04403183  - 1
cat('J12:', J12)
```


## Model 13 ##

**Create model with all variables except age and race , set cutoff at .10**
```{r}
set.seed(123)
rf13 <- randomForest(Readmitted ~ Time_hosp + Diag + Med_spec + Age_numeric +
              Gender + HbA1c + Admit + Discharge, importance = TRUE, data=train,
              cutoff = c(0.9,0.1))
```

**Model 13 training output**
**Relatively low error rate**

```{r}
rf13
```


**Make predictions using test data**
```{r}
set.seed(123)
prediction13 <-predict(rf13,test[-11])
```

**Create confusion matrix using predictions**
```{r}
cm13 = confusionMatrix(prediction4, test$Readmitted, mode = "everything")
cm13
```
**Investigate variable importances**
```{r}
var13 <- varImp(rf13)
var13
```

**Plot variable importances**
```{r}
varImpPlot(rf1)
```

**Calculate Youden's Index**
**Too low**
```{r}
cm13$byClass[1]
cm13$byClass[2]
J13 <- 0.9530637 + 0.05464191 - 1
cat('J13:', J13)
```




