---
title: "IAF 606 Final Project - Step 1: Cleaning and Data Preparation"
author: "Jeanne Reppert"
date: "4/27/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = 'C:/Users/thepi/OneDrive/Documents/UNCG/IAF 606/Final')
```

**Read file into R**

```{r}
dd <- read.csv(file="diabetic_data.csv", header=T, na.strings = "?")
```

**Load reticulate package for work in python**

```{r, echo=FALSE}
library(reticulate)
```

**Load python packages for data manipulation**

```{python, echo = FALSE}
import pandas as pd
from pandas import DataFrame
import numpy as np

```

**Read file as a pandas database into python**

```{python}
dd = pd.read_csv("diabetic_data.csv")
```

**Check data head**

```{python}
dd.head()
```

**Check number of observations and variables**

```{python}
dd.shape
```

###  Cleaning dataset in the same way that the authors of the paper cleaned the dataset  ###

**Replace '?'s and 'None's with np.nan**

```{python}
dd.replace({'?': np.nan}, regex=False,inplace=True)
dd.replace({'None': np.nan}, regex=False,inplace=True)
```

**Delete rows with discharge id's (deaths and discharges to hospice)**

```{python}
dd = dd[dd.discharge_disposition_id !=11]
dd = dd[dd.discharge_disposition_id !=13]
dd = dd[dd.discharge_disposition_id !=14]
dd = dd[dd.discharge_disposition_id !=19]
dd = dd[dd.discharge_disposition_id !=20]
dd = dd[dd.discharge_disposition_id !=21]
```

**Retain only the first encounter for each unique patient number**

```{python}
dd = dd.drop_duplicates('patient_nbr')
```

**Check number of observations and variables again, the number of observations is very close**
**to the total observations in the paper.**

```{python}
dd.shape
```

**Drop variables that were not used in the paper**

```{python}
dd = dd.drop(columns=['metformin', 'repaglinide', 'nateglinide',
                      'chlorpropamide', 'glimepiride', 
                      'acetohexamide', 'glipizide', 'glyburide', 
                      'tolbutamide','pioglitazone', 'rosiglitazone',
                      'acarbose', 'miglitol', 'troglitazone',
                      'tolazamide', 'examide', 'citoglipton', 
                      'insulin', 'glyburide-metformin',
                      'glipizide-metformin','glimepiride-pioglitazone',
                      'metformin-rosiglitazone', 'metformin-pioglitazone',
                      'admission_type_id', 'num_lab_procedures', 
                      'num_procedures', 'num_medications',
                      'number_outpatient', 'number_emergency',
                      'number_inpatient', 'diag_2', 'diag_3', 
                      'number_diagnoses', 'max_glu_serum',
                      'diabetesMed', 'payer_code', 'weight'])
```

**Check column names**

```{python}
dd.columns
```

**Rename columns for ease of display and typing**

```{python}
dd.rename(columns = {'encounter_id':'ID', 'patient_nbr':'Pat_nbr',
                     'race':'Race', 'gender':'Gender', 'age':'Age',
                     'discharge_disposition_id':'Discharge', 
                     'admission_source_id':'Admit', 
                     'time_in_hospital':'Time_hosp',
                     'medical_specialty':'Med_spec',
                     'diag_1':'Diag', 
                     'A1Cresult':'HbA1c', 'readmitted':'Readmitted'}, 
                      inplace = True)
```

###  Reorganizing variables in the same way the authors of the paper did  ###

**CLeaning HbA1c**
**Change variable HbA1c (when labeled high) based on whether meds were changed in hospital stay.**
```{python}
dd.loc[(dd['change'] == 'No') & (dd['HbA1c'] == '>8'),'HbA1c'] = 'High no change'
dd.loc[(dd['change'] == 'Ch') & (dd['HbA1c'] == '>8'),'HbA1c'] = 'High med change'

```

**After reclassifying HbA1c variable based on change variable values, removing change**
**variable from the final dataset**

```{python}
dd = dd.drop(columns=['change'])
```


**Classifying HbA1c as normal when less than 7 or labeled "Norm"**
```{python}
dd["HbA1c"].replace({"Norm": "Normal"}, inplace=True)
dd["HbA1c"].replace({">7": "Normal"}, inplace=True)

```

**Replacing NaN values with 'No test'**

```{python}
dd["HbA1c"].fillna('No test', inplace=True)
```

**Checking final value counts for HbA1c subcategories**

```{python}
a1c_count = dd['HbA1c'].value_counts() 
print(a1c_count) 
```


**Cleaning Gender**

**Dropping 'Unknown/Invalid' observations (only 3 rows)**

```{python}
dd.drop( dd[ dd['Gender'] == 'Unknown/Invalid' ].index ,
        inplace=True)
```

**Checking final value counts for Gender subcategories**

```{python}
gender_count = dd['Gender'].value_counts() 
print(gender_count) 
```


**Cleaning Readmitted**

**Reclassifying readmitted categories: readmissions after 30 days and "no" values**
**are classified as other as paper was focused on early readmissions,**
**early readmissions (readmissions <30 days) were named Readmitted**
```{python}
dd["Readmitted"].replace({"<30": "Readmitted"}, inplace=True)
dd["Readmitted"].replace({"NO": "Other", ">30":"Other"}, 
                         inplace=True)
```

**Checking final value counts for readmission subcategories**

```{python}
readmitted_count = dd['Readmitted'].value_counts()
print(readmitted_count)
```

**Cleaning Race**

**Replacing NaN values with 'missing'**

```{python}
dd["Race"].fillna('Missing', inplace=True)
```

**Consolidating Hispanic and Asian categories as "Other" and abbreviating**
**other categories**

```{python}
dd["Race"].replace({"Hispanic": "Other", "Asian": "Other", 
                  "Caucasian": "Cauc", "AfricanAmerican":"AfrAm"}, 
                   inplace=True)
```

**Checking final value counts for Race subcategories**

```{python}
race_count = dd['Race'].value_counts()
print(race_count)
```

**Cleaning Discharge**

**Replacing discharge code '1' with 'To home' and replacing other codes with 'other'**

```{python}
dd['Discharge'] = dd['Discharge'].replace([1], 'To home')
```

```{python}
dd['Discharge'] = dd['Discharge'].replace([3, 6, 18, 2, 22, 5, 25, 4, 7, 23, 
                28, 8, 15, 24, 9, 17, 10, 16, 27, 12], 'Other')
```

**Checking final value counts for Discharge subcategories**

```{python}
discharge_count = dd['Discharge'].value_counts()
print(discharge_count)
```


**Cleaning Admission**

**Replacing admission code '7' with 'From ER' and codes '1', '2', and '3' with**
**'Dr. referral', remaing codes with 'Other'**

```{python}
dd['Admit'] = dd['Admit'].replace([7], 'From ER')
```


```{python}
dd['Admit'] = dd['Admit'].replace([1, 2, 3],
                                  'Dr. referral')
```


```{python}
dd['Admit'] = dd['Admit'].replace([17, 4,6,5,20,9,8,10,22,25,14,13,11],
                                    'Other')
```

**Checking final value counts for Admission subcategories**

```{python}
admission_count = dd['Admit'].value_counts()
print(admission_count)
```

**Cleaning Medical Specialty**

**Reclassifying Medical Specialty codes and renaming them as 6 categories including**
**consolidating surgery categories and cardiology categories, categories that were not**
**part of Family/General Medicine or Internal Medicine (and were not surgery or cardiology**
**categories) are categorized as "Other", NaN values are replaced as 'unknown'**

```{python}
dd["Med_spec"].fillna('Unknown', inplace=True)
```


```{python}
dd['Med_spec'] = dd['Med_spec'].replace(['Surgery-General', 
                        'Surgery-PlasticwithinHeadandNeck', 'Surgery-Pediatric', 
                        'Surgery-Maxillofacial', 'Surgery-Colon&Rectal',
                        'SurgicalSpecialty', 'Surgery-Plastic', 'Surgeon',
                        'Surgery-Cardiovascular', 'Surgery-Thoracic', 'Surgery-Neuro', 
                        'Surgery-Cardiovascular/Thoracic', 'Surgery-General', 
                        'Surgery-Vascular'],
                        'Surgery')
```


```{python}
dd['Med_spec'] = dd['Med_spec'].replace(['Orthopedics',
            'Orthopedics-Reconstructive','Radiologist','Nephrology','Pulmonology',
            'Psychiatry','ObstetricsandGynecology','Urology','Gastroenterology',
            'Oncology','Pediatrics','PhysicalMedicineandRehabilitation',
            'Neurology','Pediatrics-Endocrinology','Otolaryngology',
            'Hematology/Oncology','Endocrinology','Pediatrics-CriticalCare',
            'Podiatry','Gynecology','Psychology','Radiology','Osteopath',
            'Hospitalist','Ophthalmology','Hematology','InfectiousDiseases',
            'Obsterics&Gynecology-GynecologicOnco','Obstetrics',
            'Anesthesiology-Pediatric','Rheumatology','OutreachServices',
            'PhysicianNotFound','Anesthesiology','Pediatrics-Neurology',
            'Endocrinology-Metabolism','Psychiatry-Child/Adolescent',
            'Pediatrics-Pulmonology','AllergyandImmunology','Pathology',
            'DCPTEAM','Dentistry','Pediatrics-EmergencyMedicine',
            'Pediatrics-Hematology-Oncology','Resident','Neurophysiology',
            'SportsMedicine','Speech','Dermatology','Perinatology',
            'Psychiatry-Addictive','Proctology','Emergency/Trauma'], 'Other')
```



```{python}
dd["Med_spec"].replace({"Cardiology-Pediatric": "Cardio", "Cardiology": "Cardio"}, 
                                inplace=True)
```


```{python}
dd["Med_spec"].replace({"Family/GeneralPractice": "Fam/GenPract"}, 
                                inplace=True)
```


```{python}
dd["Med_spec"].replace({"InternalMedicine": "IntMed"}, 
                                inplace=True)
```

**Checking final value counts for Medical Specialty subcategories**

```{python}
medical_specialty_count = dd['Med_spec'].value_counts()
print(medical_specialty_count)
```

**Cleaning Diagnosis**

**Diagnosis codes were consolidated, reclassified and renamed**
**to 9 new categories with NaN values assigned to category 'Other'**

```{python}
dd['Diag']=dd.Diag.str.replace('V+', '1000000')
```


```{python}
dd['Diag']=dd.Diag.str.replace('E+', '5000000')
```


```{python}
dd = dd.astype({"Diag": float})
```


```{python}
#1: Circulatory - 390-459, 785
#2: Digestive - 520-579, 787
#3: Genitourinary - 580-629, 788
#4: Diabetes - 250.xx
#5: Injury - 800 - 999
#6: Musculoskeletal - 710 - 739
#7: Neoplasms - 140 - 239
#8: Respiratory - 460 - 519, 786
#9: Other - 100000 - 600000
dd['Diag'] = np.where(dd['Diag'].between(390,459),
                           1000, dd['Diag'])
dd['Diag'] = dd['Diag'].replace([785], 1000)
dd['Diag'] = np.where(dd['Diag'].between(520,579), 
                           2000, dd['Diag'])
dd['Diag'] = dd['Diag'].replace([787], 2000)
dd['Diag'] = np.where(dd['Diag'].between(580,629),
                           3000, dd['Diag'])
dd['Diag'] = dd['Diag'].replace([788], 3000)
dd['Diag'] = np.where(dd['Diag'].between(250,250.99),
                           4000, dd['Diag'])
dd['Diag'] = np.where(dd['Diag'].between(800,999), 
                           5000, dd['Diag'])
dd['Diag'] = np.where(dd['Diag'].between(710,739),
                           6000, dd['Diag'])
dd['Diag'] = np.where(dd['Diag'].between(140,239), 
                           7000, dd['Diag'])
dd['Diag'] = np.where(dd['Diag'].between(460,519), 
                           8000, dd['Diag'])
dd['Diag'] = dd['Diag'].replace([786], 8000)
dd['Diag'] = np.where(dd['Diag'].between(100000000,6000000000), 
                           9000, dd['Diag'])
```


```{python}
dd['Diag'] = np.where(dd['Diag'].between(0,999), 
                           9000, dd['Diag'])
```


```{python}
dd['Diag'] = dd['Diag'].replace([1000], 'Circ')
dd['Diag'] = dd['Diag'].replace([2000], 'Dig')
dd['Diag'] = dd['Diag'].replace([3000], 'Genit')
dd['Diag'] = dd['Diag'].replace([4000], 'Diabetes')
dd['Diag'] = dd['Diag'].replace([5000], 'Injury')
dd['Diag'] = dd['Diag'].replace([6000], 'Musc/skel')
dd['Diag'] = dd['Diag'].replace([7000], 'Neoplasms')
dd['Diag'] = dd['Diag'].replace([8000], 'Resp')
dd['Diag'] = dd['Diag'].replace([9000], 'Other')
```


```{python}
dd["Diag"].fillna('Other', inplace=True)
```

**Checking final value counts for Diagnosis subcategories**

```{python}
diag_count = dd['Diag'].value_counts()
print(diag_count)
```


**Cleaning Age**

**Age is recategorized into two variables. The first takes the median age of each**
**category and assigns these observations this value as an integer, the second variable**
**reduces age to three categories (<30, 30-60, and >30).**

```{python}
dd = dd.assign(Age_numeric =lambda x: x['Age'])
```


```{python}
dd["Age_numeric"].replace({"[0-10)": 5, "[10-20)": 15, "[20-30)": 25, 
                        "[30-40)": 35, "[40-50)": 45, "[50-60)": 55,
                        "[60-70)": 65, "[70-80)": 75, "[80-90)": 85,
                         "[90-100)": 95}, inplace=True)
```

**Checking new Age_numeric variable**

```{python}
dd.Age_numeric.describe()
```


```{python}
dd["Age"].replace({"[0-10)": "<30", "[10-20)": "<30", 
                   "[20-30)": "<30", "[30-40)": "30-60",
                   "[40-50)": "30-60", "[50-60)": "30-60",
                  "[60-70)": ">60", "[70-80)": ">60",
                   "[80-90)": ">60",
                   "[90-100)": ">60"}, inplace=True)
```

**Checking final value counts for Age subcategories**

```{python}
age_count = dd['Age'].value_counts()
print(age_count)
```

**Reviewing values for Time in Hospital variable - no changes were**
**made to this variable**

```{python}
dd.Time_hosp.describe()
```


**Sending cleaned dataframe to csv file**

```{python}
dd.to_csv("cleaned_diabetic_data.csv")

```

**Loading csv file for further reconfiguring and exploration in R**

```{r}
dd <- read.csv(file="cleaned_diabetic_data.csv", header=T, na.strings = "?")
```

**Reassigning variable references**

```{r}
dd <- within(dd, Admit <- relevel(Admit, ref = 'From ER'))
dd <- within(dd, Diag <- relevel(Diag, ref = 'Diabetes'))
dd <- within(dd, Discharge <- relevel(Discharge, ref = 'To home'))
dd <- within(dd, Age <- relevel(Age, ref = '30-60'))
dd <- within(dd, HbA1c <- relevel(HbA1c, ref = 'No test'))
```

**Reviewing all variables**

```{r, echo=FALSE}
library(summarytools)
```

**Summary statistics of final cleaned Data**

```{r}

print(dfSummary(dd, graph.magnif = 0.75), method = 'render')
```

**Reading in packages in order to produce final table**

```{r, echo=FALSE}
library(sjPlot)
library(sjlabelled)
library(summarytools)
library(arsenal)
```

###  Tables to show Figure 3 information  ###
The table below was produced to match the values in table 3 in the article.

```{r, results = 'asis'}
tab <- tableby(Readmitted ~ Admit + Diag + Discharge + Time_hosp + Med_spec + Diag
                + HbA1c + Race + Gender + Age + Age_numeric, data=dd, test=FALSE)
summary(tab, text = TRUE)
```

**Writing final dataframe to csv for use in logistic regression and random**
**forest files (Parts 2 and 3)**

```{r}
write.csv(dd,"cleaned_diabetic_data.csv", row.names = FALSE)
```










































